### 2017-11-14 paper/sections/1_introduction.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

In this paper, we design cache management policies for ParSplice, driven by a
detailed analysis of the key-value accesses over the course of a long running
simulation across a variety of initial conditions. The default ParSplice
implementation uses an unlimited sized cache, shown by the ``No Cache
Management" line Figure~\ref{fig:cache-management}. This solution is
unacceptable for HPC environments, where memory is precious and a common goal
is to keep memory for such data structures below 3\%\footnote{Empirically, we
find this threshold works well for most applications}.  While users can
configure ParSplice to resize the cache when it reaches a certain threshold,
this solution requires tuning and parameter sweeps; the dashed ``Cache (too
small)" curve in Figure~\ref{fig:cache-management} shows how a poorly
configured cache hurts performance.  Our final solution, shown by the
``Dynamically Sized Cache" line in Figure~\ref{fig:cache-management}, detects
keyspace access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache
without sacrificing performance.  More importantly, it works for a variety
initial conditions without changing the policy itself.

% What is Mantle
To design more flexible cache management policies, like our ``Dynamically
Sized Cache" policy, we use the data management language and policy engine
from the Mantle paper~\cite{sevilla:sc15-mantle}. This framework allows us to
dynamically explore the effects of different software-defined cache management
strategies for the changing key-value workloads generated by ParSplice.  Mantle
was originally touted as a programmable file system metadata load balancer, but
we realize now that the collection of abstractions designed for file systems
was a control plane that improved metadata access. So in this paper we refer to
Mantle as a policy engine that injects policies written in our data management
language directly into a running service and show how this approach is useful
for reasoning about and designing different cache management strategies in
ParSplice.  By service, we mean a system that manages data and responds
requests, such as a file system or key-value store.  Developers write policies
for ``when" they want data moved and ``how much" of the data to move, then the
framework executes these policies whenever a decision needs to be made.  These
abstractions help developers unfamiliar with the domain quickly reason about,
develop, and deploy new policies that control temporal and spatial locality. We
show that Mantle:

>>>>

In this paper, we design cache management policies for ParSplice, driven by a
detailed analysis of the key-value accesses over the course of a long running
simulation across a variety of initial conditions. The default ParSplice
implementation uses an unlimited sized cache, as shown by the ``No Cache
Management" line Figure~\ref{fig:cache-management}. Using this much memory for
one cache is unacceptable for HPC environments, where memory is precious and a
common goal is to keep memory for such data structures below
3\%\footnote{Empirically, we find this threshold works well for most
applications}.  While users can configure ParSplice to resize the cache when it
reaches a certain threshold, this solution requires tuning and parameter sweeps
to find the optimal cache size that sacrifices negligible performance. The
dashed ``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows
how a poorly configured cache hurts performance.  Our final solution, shown by
the ``Dynamically Sized Cache" line in Figure~\ref{fig:cache-management},
detects keyspace access patterns and re-sizes the cache accordingly.  Without
tuning or parameter sweeps, our solution saves more memory than a hand-tuned
cache without sacrificing performance.  More importantly, it works for a
variety initial conditions without changing the policy itself.

% What is Mantle
To design more flexible cache management policies, like our ``Dynamically
Sized Cache" policy, we use the data management language and policy engine
from the Mantle paper~\cite{sevilla:sc15-mantle}. This framework allows us to
dynamically explore the effects of different software-defined cache management
strategies for the changing key-value workloads generated by ParSplice.  Mantle
was originally touted as a programmable file system metadata load balancer, but
we realize now that the collection of abstractions designed for file systems
was a control plane that improved metadata access. So in this paper we refer to
Mantle as a policy engine that injects policies written in our data management
language directly into a running service and show how this approach is useful
for reasoning about and designing different cache management strategies in
ParSplice.  By service, we mean a system that manages data and responds to
requests, such as a file system or key-value store.  Developers write policies
for ``when" they want data moved and ``how much" of the data to move, then the
framework executes these policies whenever a decision needs to be made.  These
abstractions help developers unfamiliar with the domain quickly reason about,
develop, and deploy new policies that control temporal and spatial locality. We
show that Mantle:

### 2017-11-14 paper/sections/4_dynamic-policies.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Using the Mantle policy engine, we test a variety of cache management tools and
algorithms on a single in-memory database node using the keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}. These strategies are
implemented as the ``when" and ``how much" policies from the Mantle API; the
``where" callback does not make sense for a single node (segments are either in
the cache or they are not). The evaluation metric is the accuracy and runtime
of each strategy; the strategy should be accurate enough so as to sacrifice
negligible performance and fast enough to run as often as we want to detect
regimes. The goal of the following sections is not to find an optimal solution,
as this can be done with parameter sweeps for thresholds; rather, we try to
find techniques that work for a range of inputs and system setups.

@@

  \caption{Key activity for a 4 hour run shows groups of accesses to the same
  subset of keys. Detecting these ``access regimes" leads to a more accurate cache
  management strategy.\label{fig:keyspace-zoomed}}

@@

Feeding domain-specific knowledge about the ParSplice application into a policy
leads to more a accurate cache management strategy.
Figure~\ref{fig:keyspace-zoomed} shows which keys (\(y\) axis) are accessed by
the ParSplice tasks over time (\(x\) axis). The groups of accesses to a subset
of the keys, or ``access regimes", occur because molecules are stuck in deep
trajectories. Recall that the in-memory database stores the molecules' EOM
minima, which is the smallest effective energy that a molecule observes during
its trajectory. So molecules stuck in deep trajectories explore the same minima
until they can escape to a new set of states. This exploration of the same set
of states is called a superbasin.

@@

\subsection{Failed Strategies}
These techniques proliferated more knobs that obfuscated the problem. 
\begin{itemize}
  \item Statistics
  \item Calculus
  \item K-Means
  \item DBScan
  \item Anomaly Detection
\end{itemize}

\subsection{Regime Detection}
\label{sec:regime-detection}

At each time step, we find the lowest ID and compare against the local minimum,
which is the smallest ID we have seen thus far. If we move left to right, the
local minimum never changes because the local minimum will start small. If we
move from right to left, the local minimum changes at each access regime.  For
points \(z\), \(y\), and \(x\), if the local minimum is the same we are in a
regime. Processing \(y\), we set the local minimum to be \(min(y, m_l)\), where
\(m_l\) is the local minimum of the previous time step of \(z\). The algorithm
incorrectly detects a regime change if the local minimum of \(y\) is lower than
local minimum of \(z\), since \(y\) may have points {\it within} \(z\); recall
that we are trying to detect the whole fan, not just the bottom edge of each
fan.

>>>>

Using the Mantle policy engine, we test a variety of cache management tools and
algorithms on a single in-memory database node using the keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}. These strategies are
implemented as the ``when" and ``how much" policies from the Mantle API; the
``where" callback does not make sense for a single node (segments are either in
the cache or they are not). The evaluation metric is the accuracy and runtime
of each strategy; the strategy should be accurate enough so as to sacrifice
negligible performance and fast enough to run as often as we want to detect key
access patterns. The goal of the following sections is not to find an optimal
solution, as this can be done with parameter sweeps for thresholds; rather, we
try to find techniques that work for a range of inputs and system setups.

@@

  \caption{Key activity for a 4 hour run shows groups of accesses to the same
  subset of keys. Detecting these access patterns leads to a more accurate cache
  management strategy.\label{fig:keyspace-zoomed}}

@@

Feeding domain-specific knowledge about the ParSplice application into a policy
leads to more a accurate cache management strategy.
Figure~\ref{fig:keyspace-zoomed} shows which keys (\(y\) axis) are accessed by
the ParSplice tasks over time (\(x\) axis). The groups of accesses to a subset
of keys, which we call an access ``fan" in subsequent sections, occurs because
molecules are stuck in deep trajectories. Recall that the in-memory database
stores the molecules' EOM minima, which is the smallest effective energy that a
molecule observes during its trajectory. So molecules stuck in deep
trajectories explore the same minima until they can escape to a new set of
states. This exploration of the same set of states is called a superbasin.

@@

\subsection{Failed Strategies}

To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
variety of techniques using Mantle. Unfortunately, we found that they all
proliferated more parameters that need to be tuned. Furthermore, we find that
many of the metrics do not signal a new fans, that our metrics are not accurate
enough, and that even if we tune all parameters they still will not work for
all initial conditions.  Below, we indicate which parameters we need to add for
each technique and the value we find to work best, via tuning and parameter
sweeps, for one set of initial conditions.

  \item Statistics: decay on each key counts down until 0 0-valued keys are
evicted.  ``history-of-keyspace", set to 10 seconds, to evict keys.

  \item Calculus: use derivative to strip away magnitudes; use large positive
slopes followed large negative slope as signal for new fan. ``Zero-crossing"
for distance between small/large spikes to avoid false positives, set to 40
seconds; `window size", set to 200 seconds, for the size of the moving average

  \item K-Means Clustering fails because ``K" is not known {\it a-priori} and
fans are different size. ``K", set to 4, for the number of clusters in the data
using the sum of the distances to the centroid

  \item DBScan: finds clusters using density as a metric. ``Eps", set to 20, for
max distance between 2 samples in same neighborhood; ``Min", set to 5, for the
number of samples per core

  \item Anomaly Detection: size of the image is too big and bottom edges are
not thick enough

\subsection{Dynamically Sized Cache: Access Pattern Detection}
\label{sec:regime-detection}

Figure~\ref{src:dyn-cache} shows the algorithm we use to detect the key access
patterns in Figure~\ref{fig:keyspace-zoomed}. At each time step, we find the
lowest ID and compare against the local minimum, which is the smallest ID we
have seen thus far (Line 7). We also maintain the top and bottom of each fan
(Line 13) so we can tell the ``how much" policy the number of keys to evict
(Line 19).  If we move left to right, the local minimum never changes because
the local minimum will start small. If we move from right to left, the local
minimum changes at each access regime.  For points \(z\), \(y\), and \(x\), if
the local minimum is the same we are in a regime.  Processing \(y\), we set the
local minimum to be \(min(y, m_l)\), where \(m_l\) is the local minimum of the
previous time step of \(z\).  The algorithm incorrectly detects a regime change
if the local minimum of \(y\) is lower than local minimum of \(z\), since \(y\)
may have points {\it within} \(z\); recall that we are trying to detect the
whole fan, not just the bottom edge of each fan. We exchange runtime, in the
form of an O(n) algorithm, for simplicity. This approach avoids adding new
thresholds for new fan detection, such as space between key accesses, space
between key IDs, and window size of consecutive key accesses.

\begin{figure}[h]
\footnotesize
\begin{minted}[xleftmargin=3em,linenos]{lua}
d = timeseries()          -- data (d) provided by mantle
ts, id = d:get(d:size())  -- functions provided by mantle
fan  = {start=nil, finish=ts, top=0, bottom=id}
fans = {}
for i=d:size(),1,-1 do -- detect possible fans
  ts, id = d:get(i)
  if id < fan['bottom'] then
    fan['start']  = ts
    fans[#fans+1] = fan 
    fan = {start=nil, finish=ts, top=0, bottom=id}
  end

  if id > fan['top'] then fan['top'] = id end 
end
fan['start'] = 0 
fans[#fans+1] = fan

if #fans < 2 then return false else
  WRstate(fans[#fans-1]['top']-fans[1]['bottom'])
  return true
end
\end{minted}
\caption{The ``Dynamically Sized Cache" policy from
Figure~\ref{fig:cache-management} iterates backwards over timestamp-keyID pairs
and detects when key accesses moves on to a new subset of
keys.\label{src:dyn-cache}}
\end{figure}

### 2017-11-14 paper/sections/5_file-systems.tex
paper: typos
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{General Data Management Policies}
\label{sec:scope}

@@

%- memory utilization/heuristics
From a high-level the cache management policy we designed in
Figure~\ref{src:lru-dyn} trims the cache if the cache reaches a certain size
{\it and} if it has already absorbed the initial burstiness of the workload.
Much of this implementation was inspired by the file system metadata load
balancing policy in Figure~\ref{src:lua-cephfs}, which was presented
in~\cite{sevilla:sc15-mantle}. That policy migrates load if the metadata load
is higher than the average load in the cluster {\it and} the current load has
been overloaded for more than two iterations. The two policies have the
following in common:

@@

% Load balancing in FSs
Mantle was designed for file system metadata load balancing across a cluster of
dedicated metadata servers, where spreading requests across servers improves
performance. But another technique to reduce request load is caching. If
clients and servers maintain a consistent cache, the client can do operation
locally without contacting the metadata server.  The caches in CephFS have the
same performance and utilization trade-off as ParSplice, where large caches
improve performance at the expense of a higher memory footprint. High memory
utilization is problematic when a single metadata server maintains consistent
caches with many clients .  Using cache management policies from the previous
two sections has two benefits for load balancing: (1) increases the capacity of
a single metadata server, and (2) helps identify which parts of the cache to
keep local to a server. For example, applying the regime detection algorithm from
Section~\S\ref{sec:regime-detection} to a similar file system access pattern
has the potential to identify when to migrate keys and how many keys to
migrate.

% burstiness of creates then compile
To explore this example in more detail, consider the trace of metadata requests
for compiling code in CephFS shown in Figure~\ref{fig:compile-ops}.  That trace
shows the number of file system metadata requests serviced by the metadata
server when uncompressing (\texttt{untar}), compiling (\texttt{make}), and
deleting (\texttt{rm}) the source code for the Linux kernel. If the system
knows that the job phases would progress from many creates, to many lookups, to
many deletes, then it could size its caches accordingly. For example, the file
system could cache none of the metadatas from the \texttt{untar} phase and run
regime detection during the \texttt{make} phase, resulting in the metadata
server/clients only caching metadata that is repeatedly used. For the job in
Figure~\ref{fig:compile-ops}, this would fill up the cache to only 20K inodes
(the unit of metadata for file systems) instead of 60K, resulting in almost
40MB (since an inode is about
1KB\footnote{http://docs.ceph.com/docs/master/dev/mds\_internals/data-structures/})
of memory savings without sacrificing performance.

@@

% other use cases caches/tiers 
Ceph has many other data management techniques that would benefit from the
caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\ref{sec:dom-specific}. Administrators can use the policies in ParSplice to
automatically size and manage cache tiers
\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
caching on object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain keyspace access
traces. We hypothesize that since this is all software defined caching,
something more clever than LRU would improve cache utilization and without
sacrificing too much performance.

>>>>

\section{Towards General Data Management Policies}
\label{sec:scope}

@@

%- memory utilization/heuristics
From a high-level the cache management policy we designed in
Figure~\ref{src:lru-dyn} trims the cache if the cache reaches a certain size
{\it and} if it has already absorbed the initial burstiness of the workload.
Much of this implementation was inspired by the file system metadata load
balancing policy in Figure~\ref{src:lua-cephfs}, which was presented
in~\cite{sevilla:sc15-mantle}. That policy migrates file system metadata if the
load is higher than the average load in the cluster {\it and} the current
server has been overloaded for more than two iterations. The two policies have
the following in common:

@@

% Load balancing in FSs
Mantle was designed for file system metadata load balancing across a cluster of
dedicated metadata servers, where spreading requests across servers improves
performance. But another technique to reduce request load is caching. If
clients and servers maintain a consistent cache, the client can do operations
locally without contacting the metadata server.  The caches in CephFS have the
same performance and utilization trade-off as ParSplice, where large caches
improve performance at the expense of a higher memory footprint. High memory
utilization is problematic when a single metadata server maintains consistent
caches with many clients.  Using cache management policies from the previous
two sections has two benefits for load balancing: (1) increases the capacity of
a single metadata server, and (2) helps identify which parts of the cache to
keep local to a server. For example, applying the keyspace access pattern
detection algorithm from Section~\S\ref{sec:regime-detection} to a similar file
system access pattern has the potential to identify when to migrate keys and
how many keys to migrate.

% burstiness of creates then compile
To explore this example in more detail, consider the trace of metadata requests
for compiling code in CephFS shown in Figure~\ref{fig:compile-ops}.  That trace
shows the number of file system metadata requests serviced by the metadata
server when uncompressing (\texttt{untar}), compiling (\texttt{make}), and
deleting (\texttt{rm}) the source code for the Linux kernel. If the system
knows that the job phases would progress from many creates, to many lookups, to
many deletes, then it could size its caches accordingly. For example, the file
system could cache none of the metadata from the \texttt{untar} phase and run
regime detection during the \texttt{make} phase, resulting in the metadata
server/clients only caching metadata that is repeatedly used. For the job in
Figure~\ref{fig:compile-ops}, this would fill up the cache to only 20K inodes
(the unit of metadata for file systems) instead of 60K, resulting in almost
40MB of memory savings (since an inode is about
1KB\footnote{http://docs.ceph.com/docs/master/dev/mds\_internals/data-structures/})
without sacrificing performance.

@@

% other use cases caches/tiers 
Ceph has many other data management techniques that would benefit from the
caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\ref{sec:dom-specific}. Administrators can use the policies in ParSplice to
automatically size and manage cache
tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
caching on object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain keyspace access
traces. We hypothesize that since this is all software defined caching,
something more clever than LRU would improve cache utilization without
sacrificing too much performance.
